---
title: "Renewable Energy: What factors explain the adoption of renewable energy in EU countries?"
author: Camila & Tarun
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
    orientation: rows
---

Inputs {.sidebar}
---------------------------------

Traditional energy sources like coal, oil and natural gas, have contributed to high levels of pollution and greenhouse gas emissions worldwide.

One potential way of removing the impact of greenhouse gas emission is mitigation – that is, activities intended to reduce the magnitude of climate change and its impact on human life.

The shift to cleaner forms of energy production through the adoption of renewable energy sources is a key mitigation strategy for countries like India and China.

What about the European countries? Are they also switching to renewable energy? What factors explain the adoption of renewables - wind and solar - in these countries?

Row {data-height=500 .tabset .tabset-fade}
------------------------------

###Renewable Energy Penetration
__What percentage of total electricity consumption in EU countries is generated using wind and solar energy?__  
Denmark is one of the 'greenest' country in Europe - at least when it comes to solar and wind. About 37% of its electricity needs are met through wind and solar generation. But Spain, Italy, Germany and Greece have solar and wind penetration of more than 15%, which was unthinkable even a decade ago. 

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, type = 'html', results='asis', cache=TRUE}

setwd("~/Desktop/CollaborativeDataMethods/VieiraKhanna_Assignment3/data_cleaning")

# Importing and cleaning data from Eurostat
source('Eurostat_and_Opec_data.R')

# Importing and cleaning data from World Bank
source('WDI_data.R')

# Merging the data
re_data <- merge(Combined_WDI,Combine_EuroStat,by = c("country", "year"), all.x = T, all.y = T)

# Remove missing values
re_data1 <- na.omit(re_data)

#creating var for solar and wind gen as percent of total elec consumption
re_data1$sw_pc_elec <- re_data1$re_gen/re_data1$elec_total*100 
re_data_map <- re_data1[which(re_data1$year == "2013"),]

library(googleVis)
# Plot
re_map <- gvisGeoChart(re_data_map, locationvar = 'country',
                      colorvar = 'sw_pc_elec',
                      options = list(
                        colors = "['#f7fcb9', '#31a354']",
                        region ="150"
                          ))

print(re_map, tag = 'chart')

```

###Solar Penetration

The sunny countries in __southern Europe - Italy and Greece__ - rely more on solar power than any other country. About 8% of electricity in consumed in these countries comes from solar. But __Germany and Spain__ are not far behind. In fact in terms of actual generation Germany takes the clear lead. _(see graphics below)_ 

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, type = 'html', results='asis', cache=TRUE}

#creating var for solar and wind gen as percent of total elec consumption
re_data1$s_pc_elec <- re_data1$solar_gen/re_data1$elec_total*100 
re_data_map <- re_data1[which(re_data1$year == "2013"),]

library(googleVis)
# Plot
re_map_s <- gvisGeoChart(re_data_map, locationvar = 'country',
                      colorvar = 's_pc_elec',
                      options = list(
                        colors = "['#f7fcb9', '#31a354']",
                        region ="150"
                          ))

print(re_map_s, tag = 'chart')

```

###Wind Penetration

__Denmark__ is the clear leader in wind energy. But the windy coasts of __Spain and Portugal__ are catching up fast - about 25% of electricity in these countries now comes from wind! 


```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, type = 'html', results='asis', cache=TRUE}

#creating var for solar and wind gen as percent of total elec consumption
re_data1$w_pc_elec <- re_data1$wind_gen/re_data1$elec_total*100 
re_data_map <- re_data1[which(re_data1$year == "2013"),]

library(googleVis)
# Plot
re_map_w <- gvisGeoChart(re_data_map, locationvar = 'country',
                      colorvar = 'w_pc_elec',
                      options = list(
                        colors = "['#f7fcb9', '#31a354']",
                        region ="150"
                          ))

print(re_map_w, tag = 'chart')

```

Row {data-height=550 .tabset .tabset-fade}
------------------------------

### How do different countries perform over time? 
Denmark might have the highest RE penetration but it is also small. __Germany and Spain__ are the largest markets for wind and solar power - they have high RE penetration and high absolute generation as well. 

_... See how things evolved in Europe over the last decade by clicking on the 'play' or moving the time slider located at the the bottom. You can select the factors that may impact the uptake of RE on the left side. On the right, check the countries to highlight a country._ 

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, type = 'html', results='asis'}

library(googleVis)

myState <- '\n{"colorOption":"6","time":"2013","sizeOption":"6","xZoomedDataMax":26,"xZoomedIn":false,"iconKeySettings":[{"trailStart":"2013","key":{"dim0":"Italy"}},{"trailStart":"2013","key":{"dim0":"Denmark"}},{"trailStart":"2013","key":{"dim0":"United Kingdom"}},{"trailStart":"2013","key":{"dim0":"Spain"}},{"trailStart":"2013","key":{"dim0":"Germany"}}],"xLambda":1,"xZoomedDataMin":0,"iconType":"BUBBLE","yZoomedIn":false,"yAxisOption":"11","xAxisOption":"_ALPHABETICAL","orderedByY":false,"yZoomedDataMax":37.29416288,"yZoomedDataMin":0,"showTrails":true,"uniColorForNonSelected":false,"nonSelectedAlpha":0.4,"dimensions":{"iconDimensions":["dim0"]},"playDuration":15000,"yLambda":1,"orderedByX":false,"duration":{"multiplier":1,"timeUnit":"Y"}}\n'

re_data1$year <- as.numeric(re_data1$year)

M <-  gvisMotionChart(data=re_data1, idvar="country", timevar="year", chartid="re_all_var", options=list(width=650, height=350, state=myState))

print(M,  tag = 'chart')

```

__Click on the tabs to download solar and wind generation trends (2005-13) in top 10 electricity markets in Europe__

###Wind Power Trends

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, type = 'html', results='asis', fig.height= 4, fig.width=6}

library(ggplot2)
library(dplyr)
library(plotly)

#selecting data for wind
re_data1 <- re_data1[order(re_data1$year,re_data1$wind_gen),] #arranging data
countries10 <- tail(re_data1,10)[1] %>% as.matrix() %>% as.character() %>% paste(collapse = "|") #finding top10
re_wind_top10  <- filter(re_data, grepl(countries10, country)) #selecting top 10 countries by generation 

#create graph of Wind Generation
Colours <- c('#543005','#8c510a','#bf812d','#dfc27d','#f6e8c3','#c7eae5','#80cdc1',
             '#35978f','#01665e','#003c30') # assigning colour pallette

wind_chart <- ggplot(data = na.omit(re_wind_top10)) + 
  aes(x = year, y = wind_gen, group = country, fill = country) +
  geom_area(position = "stack") + 
  scale_color_manual(values = Colours) +
  xlab("\nYear") +
  ylab("(GWh)\n") +
  ggtitle("Wind electricity in top 10 markets in the EU\n") +
  theme_bw(base_size = 10)

ggplotly(wind_chart)

```

###Solar Power Trends

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, type = 'html', results='asis', fig.height= 4, fig.width=6}

library(ggplot2)
library(dplyr)
library(plotly)

#selecting data for solar
re_data1 <- re_data1[order(re_data1$year,re_data1$solar_gen),] #arranging data
countries10 <- tail(re_data1,10)[1] %>% as.matrix() %>% as.character() %>% paste(collapse = "|") #finding top10
re_solar_top10  <- filter(re_data, grepl(countries10, country)) #selecting top 10 countries by generation 

#create graph of solar Generation
Colours <- c('#543005','#8c510a','#bf812d','#dfc27d','#f6e8c3','#c7eae5','#80cdc1',
             '#35978f','#01665e','#003c30') # assigning colour pallette

solar_chart <- ggplot(data = na.omit(re_solar_top10)) + 
  aes(x = year, y = solar_gen, group = country, fill = country) +
  geom_area(position = "stack") + 
  scale_color_manual(values = Colours) +
  xlab("\nYear") +
  ylab("(GWh)\n") +
  ggtitle("Solar electricity in top 10 markets in the EU\n") +
  theme_bw(base_size = 10)

ggplotly(solar_chart)
```


Row{data-height=300}
----------------------------

###Let there be expensive oil! 

Over the last decade as oil prices increased, it led to increased uptake of solar and wind energy in EU countries. In fact, our statistical analysis shows that high prices of oil are one of the most important variables in facilitating development of renewable energy in Europe. 

```{r, include=T, message=F, warning=FALSE, error=FALSE, echo=FALSE, type = 'html', results='asis'}

# Var. in RE generation
# Creation of a new data frame subset + sum of RE elect. generation in each year for all countries
re_data2 <- aggregate(re_data1$re_gen, list(re_data1$year), FUN=sum)
names(re_data2)[names(re_data2) == 'x'] <- 're_gen_sum'
names(re_data2)[names(re_data2) == 'Group.1'] <- 'year'

#Percentage changes in RE generation from 2005 until 2013
re_data2$re_gen_pct <- c(NA, 17.93, 27.53, 17.38, 15.80, 16.80, 30.69, 21.54, 16.15)

#Var. in Oil prices
re_data2$oil_pct <- c(NA, 20.61, 13.09, 36.72, -35.35, 26.84, 38.74, 1.85, -3.27)

library(googleVis)
#Plotting graph
Dashed <-  gvisLineChart(re_data2, xvar = 'year', yvar=c('re_gen_pct','oil_pct'),
                         options=list(series="[{color:'green', targetAxisIndex: 0, 
                         lineWidth: 1, lineDashStyle: [2, 2, 20, 2, 20, 2]}, 
                         {color: 'blue',targetAxisIndex: 1, 
                         lineWidth: 2, lineDashStyle: [4, 1]}]",
                                 vAxes="[{title:'% change in RE Generation'}, {title:'% change in oil prices'}]"
                         ))

print(Dashed, tag = 'chart')
```

